---
title: "data wrangling i session 2"
output: github_document
date: "2023-09-21"
---


```{r}
library(tidyverse)
litters_df=
  read_csv("data/fas_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df =
  read_csv("data/fas_pups.csv")
pups_df = janitor::clean_names(pups_df)


```


##select function

"select" is used to select columns (variable)
(will put them in the order you right them)
":" is used to say "everything between this variable and that variable"
starts_with can be used to select columns that start with a certain term
Use - to take out a specific variable
"everything" says I want to keep all the variables, but lets you reorder your columns.

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, group, starts_with("pups"))

select(litters_df, -litter_number)

select(litters_df, -starts_with("gd"))

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())

select(litters_df, gd0_weight, everything())

```

select vs. pull
```{r}
select(litters_df, group)
pull(litters_df, group)
```

