---
title: "data wrangling i session 2"
output: github_document
date: "2023-09-21"
---


```{r}
library(tidyverse)
litters_df=
  read_csv("data/fas_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df =
  read_csv("data/fas_pups.csv")
pups_df = janitor::clean_names(pups_df)


```


##select function

"select" is used to select columns (variable)
(will put them in the order you right them)
":" is used to say "everything between this variable and that variable"
starts_with can be used to select columns that start with a certain term
Use - to take out a specific variable
"everything" says I want to keep all the variables, but lets you reorder your columns.

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, group, starts_with("pups"))

select(litters_df, -litter_number)

select(litters_df, -starts_with("gd"))

select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())

select(litters_df, gd0_weight, everything())

```

```{r}
rename(litters_df, litter_id = litter_number)
relocate(litters_df, litter_number)
```

select vs. pull
```{r}
select(litters_df, group)
pull(litters_df, group)
```


learning assessment
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

##"filter"

get rid of rows using filter()
== tells you what's true that should be kept

```{r}

filter(litters_df, group == "mod8")

filter(litters_df, group != "Con7")

filter(litters_df, gd0_weight > 20)

filter(litters_df, gd0_weight <= 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df !(group=="Cony"))

filter(litters_df, group %in% c("Con7", "Con8", gd0_weight > 20))
```


##mutate

this is used to add or change variables. In the example below, a new variable called weight gain is comprised of weight 18 - weight 0

```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)
mutate(litters_df, group = str_to_lower(group))

mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group),
)

```



what about missing rows?
drop_na drops any missing values that = NA in the dataset. you can also specify which column to delete from.

```{r}
drop_na(litters_df, gd0_weight)
```


##arange

helpful for nested datasets

```{r}
arrange(litters_df, gd0_weight)
arrange(litters_df, group, gd0_weight)
view(arrange(litters_df, group, gd0_weight))
```


## pipes
shift command m creates |>
Below does the same as all the code we wrote above, just cleaner and more legible

```{r}
litter_df =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(-starts_with("pups")) |> 
  mutate(
    group = str_to_lower(group),
    wt_gain = gd18_weight - gd0_weight,
  ) |> 
  drop_na(wt_gain) |> 
  arrange(group, wt_gain)

```

