---
title: "Session 7"
output: html_document
date: "2023-09-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

##pulse dataset

haven helps us read data, specifically read_sas helps with sas files

in the original dataset, the bdi score and each visit was spread across several variables. below changes from y format to long format, so that all those variables are values of one combined variable (visit) and then bdi_score

will do extra tidying to make sure naming convention is consistent across bdi_scores

in a mutate statement, replace is used to change values in one variable to another value
```{r}
pulse_df =
  haven::read_sas("data/public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "bdi_score_"
  ) |> 
  mutate(
    visit = replace(visit, visit == "bl", "00m")
  )


pulse_df
```


##learning assessment

```{r}

litters_df =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |> 
  select(litter_number, gd0_weight, gd18_weight) |> 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight",
  )  |> 
  mutate(
    gd = case_match(
      gd, 
      "0_weight" ~ 0,
      "18_weight" ~ 18,
  )

litters_df
```

##bind rows LOTR

Import LOTR words data

read function asks which columns you want to read in from an xcl file

any time you copy and paste code, you are at risk of making mistakes. you should always try to write a function instead (which we will do later!)

remember it is case sensitive! after you use "clean names" it makes everything the same case, so if you have later in your code "Male" it will say that column doesn exist.

```{r}
fellowship_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "B3:D6") |> 
  mutate(movie = "fellowship")


two_towers_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "F3:H6") |> 
  mutate(movie = "two towers")

return_of_king_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "J3:L6") |> 
  mutate(movie = "return of the king")

lotr_df =
  bind_rows(fellowship_df, two_towers_df, return_of_king_df) |>
  janitor::clean_names() |> 
  pivot_longer(
    male:female,
    names_to = "gender",
    values_to = "word"
  ) |> 
  relocate(movie)

lotr_df
```



